Association between SARS-CoV-2 and metagenomic content of samples from the Huanan Seafood Market

### 2023-06-29 paper/paper.tex
fix typo in methods (#3)
Methods incorrectly said raccoon dog mitochondrial genome was manually downloaded when it was actually Amur hedgehog. Code is correct, typo is just in methods.

See https://twitter.com/BenFPiercePhD/status/1654364200209207298

Furthermore, the correlations are highly contingent on the sample set and details of how the statistics are calculated.
For instance, if we exclude the first sampling timepoint and only look at samples calculated on January-12-2020 or later, then the most correlated species is the Oriental rat snake, with the correlation for humans becoming even weaker and that for raccoon dogs remaining negative (Figure~\ref{fig:per_species_corr_faceted_later_dates}).
The correlations also change if only consider samples containing a non-zero number of SARS-CoV-2 reads, or calculate correlations on a linear rather than log scale (Figure~\ref{fig:different_corrs}).
The interactive versions of the correlation plots at \url{https://jbloom.github.io/Huanan_market_samples/per_species_corr_faceted.html} and \url{https://jbloom.github.io/Huanan_market_samples/per_species_corr_single.html} enable the reader to explore the correlations for different sample subsets and methods of calculating the statistics.
Exploration of these interactive plots shows that it takes a great deal of cherry picking of parameter combinations to put any plausibly infectable hosts at the top of the list of species whose mitochondrial material is most correlated with the abundance of SARS-CoV-2.

@@

To create an mitochondrial genome alignment reference set, I followed a procedure partially analogous to that described by \citet{crits2023genetic}.
All mitochondrial genomes in the RefSeq database were downloaded, and then filtered to retain only genomes from the phylum Chordata (in contrast, \citet{crits2023genetic} describe retaining all metazoa mitochondrial genomes).
The reason I limited to chordate rather than metazoa mitochondrial genomes is that all chordates have very similar length mitochondrial genomes, but at the level of metazoa there is wide variation in mitochondrial genome length.
The raccoon dog mitochondrial genome was not present in this set, so it was separately downloaded from Genbank accession KX964606.
All the mitochondrial genomes were then filtered to remove highly similar ones.
To do this, \texttt{Mash}~\citep{ondov2016mash} was used to compute the mash distances between all pairs of mitochondrial genomes.
To ensure the most relevant mitochondrial genomes were retained, I first specified for manual retention the genomes for all the species listed in the third supplementary file of \citet{crits2023genetic}, as well as genomes for some additional relevant species (these species are listed under the \textit{mitochondrial\_genomes\_to\_keep} key in \url{https://github.com/jbloom/Huanan_market_samples/blob/main/config.yaml}).
After retaining all of these manually specified mitochondrial genomes, the pipeline then greedily iterated through all other genomes choosing for retention each genome that was not within a mash distance of 0.07 of another already retained genome.
The full set of retained mitochondrial genomes is listed in Table~\ref{tab:mito_retained}.

>>>>

Furthermore, the correlations are highly contingent on the sample set and details of how the statistics are calculated.
For instance, if we exclude the first sampling timepoint and only look at samples calculated on January-12-2020 or later, then the most correlated species is the Oriental rat snake, with the correlation for humans becoming even weaker and that for raccoon dogs remaining negative (Figure~\ref{fig:per_species_corr_faceted_later_dates}).
The correlations also change if only consider samples containing a non-zero number of SARS-CoV-2 reads, or calculate correlations on a linear rather than log scale (Figure~\ref{fig:different_corrs}).
The interactive versions of the correlation plots at \url{https://jbloom.github.io/Huanan_market_samples/per_species_corr_faceted.html} and \url{https://jbloom.github.io/Huanan_market_samples/per_species_corr_single.html} enable the reader to explore the correlations for different sample subsets and methods of calculating the statistics.
%Exploration of these interactive plots shows that it takes a great deal of cherry picking of parameter combinations to put any plausibly infectable hosts at the top of the list of species whose mitochondrial material is most correlated with the abundance of SARS-CoV-2.

@@

To create an mitochondrial genome alignment reference set, I followed a procedure partially analogous to that described by \citet{crits2023genetic}.
All mitochondrial genomes in the RefSeq database were downloaded, and then filtered to retain only genomes from the phylum Chordata (in contrast, \citet{crits2023genetic} describe retaining all metazoa mitochondrial genomes).
The reason I limited to chordate rather than metazoa mitochondrial genomes is that all chordates have very similar length mitochondrial genomes, but at the level of metazoa there is wide variation in mitochondrial genome length.
The Amur hedgehog mitochondrial genome was not present in this set, so it was separately downloaded from Genbank accession KX964606.
All the mitochondrial genomes were then filtered to remove highly similar ones.
To do this, \texttt{Mash}~\citep{ondov2016mash} was used to compute the mash distances between all pairs of mitochondrial genomes.
To ensure the most relevant mitochondrial genomes were retained, I first specified for manual retention the genomes for all the species listed in the third supplementary file of \citet{crits2023genetic}, as well as genomes for some additional relevant species (these species are listed under the \textit{mitochondrial\_genomes\_to\_keep} key in \url{https://github.com/jbloom/Huanan_market_samples/blob/main/config.yaml}).
After retaining all of these manually specified mitochondrial genomes, the pipeline then greedily iterated through all other genomes choosing for retention each genome that was not within a mash distance of 0.07 of another already retained genome.
The full set of retained mitochondrial genomes is listed in Table~\ref{tab:mito_retained}.
